
---
title: "Peer assessment 1"
author: "Gil TOSSOU"
date: "May 15, 2016"
output: 
  html_document
---
# Reproducible Research: Peer Assessment 1

## Load the data from working directory into data frame
```{r}
activity<-read.csv(file = "activity.csv",header = TRUE)
```
### Quick look to data (first 7 rows) and dimensionof the data
```{r}
dim(activity)
head(activity,6)
```
###check data types
```{r}
str(activity)
```
We notice that the *date* variable is a *factor* variable. 

### Let's convert it into *date* object
```{r}
activity$date=as.Date(activity$date)
```
## Mean and median number of steps taken each day
```{r}
sumStepsDate=tapply(activity$steps,activity$date,sum)
summary(sumStepsDate)
```

## Histogram of the total number of steps taken each day
```{r totalStepsByDay, fig.width=8,fig.height=4,echo=TRUE}
sumStepsDateNonas<-sumStepsDate
hist(sumStepsDate)
```
## The 5-minute interval that, on average, contains the maximum number of steps
 ```{r}
sumStepsDate <- data.frame(template=names(sumStepsDate),mean=sumStepsDate)
colnames(sumStepsDate)<-c("date","steps")
row.names(sumStepsDate) <- NULL
```
### Get date feature to date type and remove nas
 ```{r}
sumStepsDate$date<-as.Date(sumStepsDate$date)
sumStepsDate<-na.omit(sumStepsDate)
```
### Now, the plot
```{r}
plot(sumStepsDate$date,sumStepsDate$steps,type="l",col="blue")
```
##  The 5-minute interval that, on average, contains the maximum number of steps
```{r}
factorActivity<-activity
factorActivity$interval<-as.factor(factorActivity$interval)
meanByInterval<-tapply(factorActivity$steps,factorActivity$interval,mean,na.rm=T)

summary(meanByInterval)
```
## Code to describe and show a strategy for imputing missing data
### Missing values by variables
```{r}
table(is.na(activity$steps))
table(is.na(activity$interval))
table(is.na(activity$date))
```
### The nas values are only in the *steps* variable

### Checking for a pattern in the missing values
```{r}
table(activity$date,is.na(activity$steps))
```
### Let's just remove NAs. To do : use mice or impute() to impute the missing values 
```{r}
activity_clean=na.omit(activity)
```
## Histogram of the total number of steps taken each day after missing values are imputed
```{r totalStepsByDayNonas, fig.width=8,fig.height=4,echo=TRUE}
hist(sumStepsDateNonas)
```
